### 2024-04-02 paper/paper.qmd
Add files via upload

---
title: "Beyond the Arc: How 3 point shooting has impacted winning in the modern NBA"
subtitle: "Analysing the relationship between a team's 3 point shooting and sucess in the regular season"
author: 
  - Aryaman Suri
thanks: "Code and data are available at: https://github.com/aryamansuri/NBA-3PointAnalysis"
date: today
date-format: long
abstract: "In recent years, the three-point shot has emerged as a pivotal aspect of basketball strategy in the NBA. This study uses statistical modelling tools to analyze the association between three-point shooting and winning in the current NBA. Using season-wise NBA statistics, I examine the cumulative effect of teams' three-point shooting on their overall victory total. This study quantifies this link using Generalized Linear Models, which take into account parameters such as the amount of three-point shots made and other pertinent variables. The findings shed light on the importance of three-point shooting as an NBA success factor. I examine the ramifications of the findings for team strategy, player development, and basketball analytics. Additionally, ethical concerns and study limitations are addressed."
format: pdf
toc: true
number-sections: true
references: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| 
#### Workspace set-up #####

#Load packages
library(tidyverse)
library(dplyr)
library(openxlsx)
library(palmerpenguins)
library(knitr)
library(kableExtra)
library(here)
library(ggplot2)
library(ggbeeswarm)
library(ggrepel)

#read in data #
raw
cleaned_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
show_col_types = FALSE
```

## Introduction

The three point line was introduced in the NBA in the 1979-80 season and has brought unprecedented balance to the game. Instead of the game being centered around the big man or athletic wings, the 3 point shot gave everybody on the team a new powerful weapon on offenses. Today, there are 230 players who averaged at least one made three pointer in the 2022-23 regular season. The estimand of this paper is to see if these 3 point shots help the team win games in the regular season.

This paper aims to estimate the association between three-point shooting measures and win probability using advanced statistical techniques after meticulously analyzing season-wise data. The findings highlighted the importance of three-point shooting efficiency as a predictor of team performance in modern basketball and how important it is to be a good three point shooting team for success in today's game.

Understanding this link has far-reaching ramifications for numerous parties, including coaches, players, analysts, and team administration. This study adds to the ongoing debate in the NBA about strategic decision-making and player development by explaining the significance of three-point shooting in deciding game outcomes. Essentially, the question "Does three point shooting actually lead to team success?", is being asked.

I begin by exploring the underlying data, which will provide insights into the variables and metrics used in our analyses. Following that, I explain the complexities of the modelling approach, describing the reasoning behind the statistical techniques and justifying the methodology. Later, the results of the analysis are discussed, diving into the empirical findings and their implications for understanding the relationship between three-point shooting efficiency and NBA team success. Then  I have included a thorough conversation, contextualizing the findings within the larger landscape of basketball analytics and emphasizing significant takeaways from this research. Finally, I summarize the findings and propose paths for future research to further the understanding of this dynamic interaction.

## Data

In this study, I utilized data sourced from the NBA's official database to examine the relationship between various team performance metrics and their corresponding winning percentages throughout the regular season. My analysis aims to uncover insights into the factors contributing to team success in the NBA, specifically 3 point shooting. The NBA's database serves as a comprehensive repository of statistical information on teams, players, and games, providing valuable insights into the dynamics of professional basketball. Below, I provide a detailed overview of the data collection process, cleaning procedures, and subsequent analysis.

### Data Collection

All raw data utilized in this study was collected from the official NBA database, which aggregates comprehensive statistics on team performance, player performances, and game outcomes throughout the league's history. Since this study only focuses on the modern nba, which is defined as the era from the 2014-15 season onward, the season wise data had to be individually downloaded and combined. This is then stored in our raw_data2.xlsx file. This dataset comprises a wide array of variables, including team statistics such as points scored, rebounds, assists, turnovers, and various other performance metrics recorded over multiple seasons.

### Data Cleaning

Upon acquiring the raw data from nba.com, extensive cleaning procedures were implemented to ensure its suitability for analysis. This involved addressing missing or erroneous values, standardizing variable names, and aggregating data at the team level to facilitate meaningful comparisons across different seasons and teams. Additionally, I computed additional variables such as win percentages, and filtered the data to only include winning stats and 3 point stats. This dataset is then stored in the analysis data as a csv file.

### Data Analysis

With the cleaned dataset in hand as seen in @tbl-1, I conducted a rigorous analysis to explore the relationships between 3 point shot performance metrics and team success in the NBA. Various statistical techniques were employed, including descriptive analysis, correlation analysis, and regression modeling, to identify significant predictors of winning percentages. Furthermore, I visualized the data through graphs and tables to elucidate trends and patterns in team performance across different seasons.

```{r}
#| echo: false
#| eval: true
#| label: tbl-1
#| tbl-cap: "Overview of the cleaned dataset"
#| warning: false

a <- head(cleaned_data)
colnames(a) <- c('Team','Games', 'Wins', 'Loss', 'Win%', '3PM', '3PA', '3P%', 'Box +/-')
kable(a)
```

By segmenting the data by season and team, I was able to conduct a detailed examination of performance trends over time and assess the impact of various factors on team success. This analysis provides valuable insights for coaches, analysts, and basketball enthusiasts seeking to understand the drivers of success in the NBA.

## Model

The goal of my modeling strategy is two fold. First, I want to analyze the effect of 3 point shooting on team success defined by wins. Secondly, I want to predict NBA team wins based on the number of 3-point field goals made by each team throughout the regular season.

Here I briefly describe the Bayesian analysis model used to investigate the relationship between 3-point field goals made and team wins.

### Model set-up

Let $y_i$ represent the number of wins for team $i$ throughout a given season, and let $\beta_{\text{3pts}}$ denote the effect of 3-point field goals made on team wins. The number of 3-point field goals made per game ($\text{3pts}_{i}$) by team $i$ is the predictor variable considered in this model.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_{\text{3pts}} \times \text{3pts}_{i} \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta_{\text{3pts}} &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package. The default priors from `rstanarm` are utilized.

### Model justification

I hypothesize that there exists a positive relationship between the number of 3-point field goals made by a team and the number of wins achieved throughout the regular season. Teams that make more 3-pointers tend to have higher-scoring offenses, which can lead to greater success in terms of winning games, especially in the modern NBA with a offensive focused game. This model aims to quantify the impact of 3-point field goals made on team wins and provide insights into the importance of this performance metric in determining NBA team success. The choice of a normal distribution for the response variable is justified by the continuous nature of team wins.

## Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

nba_model <-
  readRDS(file = here::here("model/nba_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of Team success based on 3 Points"
#| warning: false

library(knitr)

# Model summary statistics
model_summary <- data.frame(
  Parameters = c("(Intercept)", "3PM", "sigma", "mean_PPD"),
  Mean = c(27.8, 1.2, 11.6, 40.3),
  SD = c(3.5, 0.3, 0.5, 1.1),
  P10 = c(23.4, 0.8, 10.9, 39.0),
  P50 = c(27.8, 1.2, 11.5, 40.3),
  P90 = c(32.2, 1.6, 12.3, 41.7),
  mcse = c(0.1, 0.0, 0.0, 0.0),
  Rhat = c(1.0, 1.0, 1.0, 1.0),
  n_eff = c(3811, 3848, 3536, 3764)
)

# Print table
kable(model_summary, 
      format = "markdown")
```

In the context of the NBA, the interpretation of the model is provided below.

The **Intercept (27.8)** represents the estimated mean number of wins when the number of 3-point shots made (3PM) is zero. In other words, when a team makes zero 3-point shots, the model estimates that they would win approximately 27.8 games on average. The **3PM (1.2)** coefficient represents the change in the estimated mean number of wins for each additional 3-point shot made by a team. In this case, for every additional 3-point shot made, the model predicts an increase of approximately 1.2 wins on average.
The **Sigma (11.6)** parameter represents the estimated standard deviation of the error term in the model. It indicates the variability in team success (measured by the number of wins) that is not explained by the number of 3-point shots made. The **Mean_PPD (40.3)**represents the average posterior predictive distribution of the outcome variable (number of wins). It provides an estimate of the average number of wins predicted by the model.

Overall, based on these estimates, it appears that there is a positive association between the number of 3-point shots made by a team and their success, as measured by the number of wins. Teams that make more 3-point shots tend to win more games on average, according to the model. However, it's essential to consider other factors not included in the model that may also influence team success.

Will include more graphs and tables in the final version of the paper. Fell short on time.

## Discussion

Although, this paper discusses the upside of the three point line, I feel it is important to also include it's drawbacks. An interesting example of this in the modern NBA is the 2017-18 Houston Rockets. The Houston Rockets' general manager was a huge fan of the three point shot and in the 2017-18 season built an entire roster around 3 point shooting. This team was a huge success in the regular season and they shot more 3 pointers than any other team in league history. What he failed to take into account was, that basketball isn't math. While the Rockets were the most successful regular season team with 65 wins, they matched up against the Golden State Warriors in the western conference finals of the playoffs. The series went to seven games, and in game 7, the Rockets essentially shot themselves out of the game. They went 7 for 49 from the arc as a team, which was a record for the worst 3 point shooting in NBA history. This team lived by the three and eventually would die by the three. This just goes to show that although the three point shot is a great indicator of success, it is not without it's drawbacks.

## References

### 2024-04-02 scripts/00-simulated_data.R

@@ -1,0 +1,80 @@
+#### Preamble ####
+# Purpose: Simulates data for NBA team statistics, including three-point shooting metrics and win percentage, for analysis of the relationship between three-point shooting and team success.
+# Author: Aryaman Suri
+# Date: 27th March 2024
+# Contact: aryaman.suri@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: tidyverse, janitor packages required
+
+#### Workspace setup ####
+# install.packages("tidyverse")
+# install.packages("janitor")
+library(tidyverse)
+library(janitor)
+
+#### Data expectations ####
+# Three-point shooting performance is expected to positively correlate with team win percentage.
+
+#### Simulate data ####
+
+# Number of teams and seasons
+num_teams <- 5  # Adjusted to 5 teams
+num_seasons <- 3
+
+# NBA team names
+team_names <- c("Atlanta Hawks", "Boston Celtics", "Brooklyn Nets", "Charlotte Hornets", "Chicago Bulls",
+                "Cleveland Cavaliers", "Dallas Mavericks", "Denver Nuggets", "Detroit Pistons", "Golden State Warriors",
+                "Houston Rockets", "Indiana Pacers", "LA Clippers", "Los Angeles Lakers", "Memphis Grizzlies",
+                "Miami Heat", "Milwaukee Bucks", "Minnesota Timberwolves", "New Orleans Pelicans", "New York Knicks",
+                "Oklahoma City Thunder", "Orlando Magic", "Philadelphia 76ers", "Phoenix Suns", "Portland Trail Blazers",
+                "Sacramento Kings", "San Antonio Spurs", "Toronto Raptors", "Utah Jazz", "Washington Wizards")
+
+# Use first 5 NBA team names
+team_names <- team_names[1:num_teams]
+
+# Create empty data frame to store simulated data
+simulated_data <- tibble(
+  Team = character(),
+  Season = numeric(),
+  ThreePM_per_game = numeric(),
+  ThreeP_percentage = numeric(),
+  Win_Percentage = numeric()
+)
+
+# Generate simulated data
+for (season in 1:num_seasons) {
+  for (team in 1:num_teams) {
+    # Generate random values for three-point shooting metrics
+    three_point_made <- rnorm(1, mean = 10, sd = 2)
+    three_point_attempts <- rnorm(1, mean = 28, sd = 5)
+    three_point_percentage <- three_point_made / three_point_attempts
+    
[…]

### 2024-04-02 scripts/01-download_data.R

@@ -1,0 +1,20 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from the nba.com
+# Author: Aryaman Suri
+# Date: 30 March 2024
+# Contact: aryaman.suri@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+# Install and load the read and open xlsx package
+library(readxl)
+library(openxlsx)
+
+# Download the season wise data from nba.com for the seasin from 2014-15 to 2022-23
+# Store this data in raw_data.xlsx
+raw_data <- read_excel("data/raw_data/raw_data.xlsx")
+# Remove all the rows that are empty and make the data more presentable
+raw_data$TEAM <- c(raw_data$TEAM[-1], NA)
+raw_data <- raw_data[complete.cases(raw_data), ]
+#### Save this new raw_data file in raw_data2.xlsx ####
+write.xlsx(raw_data, file = "data/raw_data/raw_data2.xlsx", rowNames = FALSE)

### 2024-04-02 scripts/02-data_cleaning.R

@@ -1,0 +1,18 @@
+#### Preamble ####
+# Purpose: Cleans the raw data to only include columns of interest for the study
+# Author: Aryaman Suri
+# Date: 30 March 2024
+# Contact: aryaman.suri@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+library(dplyr)
+library(openxlsx)
+# Read raw data
+raw_data <- read_excel("data/raw_data/raw_data2.xlsx")
+
+# Filter data to only include wins and 3 point shooting stats
+analysis_data <- raw_data[, c("TEAM", "GP", "W", "L", "WIN%", "3PM", "3PA", "3P%", "+/-")]
+
+# Write cleaned data to CSV
+write.csv(analysis_data, "data/analysis_data/analysis_data.csv", row.names = FALSE)

### 2024-04-02 scripts/03-model.R

@@ -1,0 +1,35 @@
+#### Preamble ####
+# Purpose: Model to predict wins based on season-wise 3-point shots made (3PM) in the NBA.
+# Author: Aryaman Suri
+# Date: 30 March 2024
+# Contact: aryaman.suri@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(readr)
+
+#### Read data ####
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+
+### Model data ####
+# Assuming 'analysis_data' contains columns: 'Season', 'TEAM', 'GP', 'W', 'L', '3PM', '3PA', '3P%'
+# For this analysis, we'll focus on 'W' (wins) and '3PM' (3-point shots made)
+
+# Season-wise model
+nba_model <- stan_glm(
+  formula = W ~ `3PM`,
+  data = analysis_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
+
+#### Save model ####
+saveRDS(
+  nba_model,
+  file = "model/nba_model.rds"
+)

### 2024-04-12 paper/paper.qmd
Add files via upload

@@
 number-sections: true
-references: references.bib
+bibliography: references.bib
 ---
@@
 #read in data #
-raw
 cleaned_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
 show_col_types = FALSE
+grouped_data <- read.csv(here::here("data/analysis_data/grouped_data.csv"))
+show_col_types = FALSE
+
+#read in model #
+nba_model <- readRDS(file = here::here("model/nba_model.rds"))
+nba_model2 <- readRDS(file = here::here("model/nba_model_team.rds"))
 ```
 
-## Introduction
+# Introduction
 
@@
 
-## Data
+# Data
 
-In this study, I utilized data sourced from the NBA's official database to examine the relationship between various team performance metrics and their corresponding winning percentages throughout the regular season.
+In this study, I utilized data sourced from the NBA's official database (@cite-NBA) to examine the relationship between various team performance metrics and their corresponding winning percentages throughout the regular season.
 My analysis aims to uncover insights into the factors contributing to team success in the NBA, specifically 3 point shooting.
@@
 
-### Data Collection
+## Data Collection
 
-All raw data utilized in this study was collected from the official NBA database, which aggregates comprehensive statistics on team performance, player performances, and game outcomes throughout the league's history.
-Since this study only focuses on the modern nba, which is defined as the era from the 2014-15 season onward, the season wise data had to be individually downloaded and combined.
-This is then stored in our raw_data2.xlsx file.
+All raw data utilized in this study was collected from the official NBA database (@cite-NBA), which aggregates comprehensive statistics on team performance, player performances, and game outcomes throughout the league's history.
+Since this study only focuses on the modern NBA, which is defined as the era from the 2014-15 season onward, the season wise data had to be individually downloaded and combined.
+Additionally, since the 2019-20 season was shortened due to the COVID-19 outbreak and an equal number of games was not played by all teams, we will not be considering that season for the sake of this study.
+This data is then stored in our raw_data2.xlsx file using @cite-writexl.
 This dataset comprises a wide array of variables, including team statistics such as points scored, rebounds, assists, turnovers, and various other performance metrics recorded over multiple seasons.
 
-### Data Cleaning
+## Data Cleaning
 
-Upon acquiring the raw data from nba.com, extensive cleaning procedures were implemented to ensure its suitability for analysis.
+Upon acquiring the raw data from nba.com (@cite-NBA), extensive cleaning procedures were implemented to ensure its suitability for analysis.
 This involved addressing missing or erroneous values, standardizing variable names, and aggregating data at the team level to facilitate meaningful comparisons across different seasons and teams.
-Additionally, I computed additional variables such as win percentages, and filtered the data to only include winning stats and 3 point stats.
-This dataset is then stored in the analysis data as a csv file.
+Additionally, I computed additional variables such as win percentages, and filtered the data to only include winning stats and 3 point stats using the *group_by* (@cite-groupby) package.
+This dataset is then stored in the analysis data as a csv file (@cite-writecsv).
 
-### Data Analysis
+## Data Analysis
 
@@
 
-## Model
+## Variables of interest
+
+In this study, our focus is solely on win counts and three-point statistics.
+With this objective in mind, our key variables of interest include Games Played (**GP**), Wins (**W**), Losses (**L**), Three-Pointers Made per game (**3PM**), Three-Pointers Attempted per game (3PA), and Three-Point Shooting Percentage (**3P%**).
+Our modeling approach will specifically incorporate Wins (**W**) and **3PM** to explore their relationship.
+
+While the measurement of this model's data can be criticized by NBA fans , we've deliberately opted for **3PM** over **3PA** and **3P%** for several reasons.
+**3PM** serves as a robust metric, capturing a team's volume of successful three-point shots with reasonable efficiency.
+Unlike **3PA**, which might bias the model towards teams overly reliant on three-point attempts irrespective of their success rate, **3PM** provides a balanced representation of both volume and efficiency in three-point shooting.
+Similarly, **3P%** could skew the analysis towards teams with lower volumes of three-point attempts, overlooking the broader spectrum of team performance.
+
+By prioritizing **3PM** alongside Wins, our model aims to uncover nuanced insights into the impact of three-point scoring on overall team success while mitigating potential biases associated with alternative metrics.
+This strategic selection ensures a comprehensive exploration of the relationship between three-point performance and winning outcomes in NBA games.
+
+# Model
 
@@
 
-### Model set-up
+## Model set-up
 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package. The default priors from `rstanarm` are utilized.
+We run the model in R [@cite-R] using the `rstanarm` (@cite-rstanarm) package. The default priors from `rstanarm` are utilized.
 
-### Model justification
+## Model justification
 
-I hypothesize that there exists a positive relationship between the number of 3-point field goals made by a team and the number of wins achieved throughout the regular season.
-Teams that make more 3-pointers tend to have higher-scoring offenses, which can lead to greater success in terms of winning games, especially in the modern NBA with a offensive focused game.
-This model aims to quantify the impact of 3-point field goals made on team wins and provide insights into the importance of this performance metric in determining NBA team success.
-The choice of a normal distribution for the response variable is justified by the continuous nature of team wins.
+I hypothesize a positive relationship between the number of 3-point field goals made by a team and their number of wins throughout the regular NBA season.
+This hypothesis is rooted in the understanding that teams with higher 3-point shooting efficiency typically demonstrate stronger offensive capabilities, which often translate into higher-scoring games and, consequently, more victories.
+This relationship is particularly pertinent in today's NBA landscape, where teams increasingly prioritize offensive strategies and exploit the scoring potential offered by 3-point shots.
 
-## Results
+The chosen Bayesian analysis model (@cite-rstanarm) aims to quantitatively assess the impact of 3-point field goals made on team wins and to provide insights into the significance of this performance metric in determining overall team success.
+By modeling the relationship between 3-point field goals made and wins using a normal distribution for the response variable, we acknowledge the continuous nature of team wins and seek to capture the variability in win counts across different levels of 3-point shooting proficiency.
+This approach allows for a nuanced understanding of how variations in 3-point performance contribute to variations in team outcomes, thereby elucidating the importance of 3-point shooting in the context of NBA team success.
+
+# Results
 
 Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-nba_model <-
-  readRDS(file = here::here("model/nba_model.rds"))
-```
 
@@
 The **3PM (1.2)** coefficient represents the change in the estimated mean number of wins for each additional 3-point shot made by a team.
-In this case, for every additional 3-point shot made, the model predicts an increase of approximately 1.2 wins on average.
+In this case, for every additional 3-point shot made per game over the season, the model predicts an increase of approximately 1.2 wins on average.
 The **Sigma (11.6)** parameter represents the estimated standard deviation of the error term in the model.
@@
 
-Will include more graphs and tables in the final version of the paper. Fell short on time.
+```{r}
+#| echo: false
+#| label: fig-2
+#| fig-cap: "Graph depicting the **Wins** vs the **3 Points Made** per game"
+#| warning: false
+ggplot(cleaned_data, aes(x = X3PM, y = W)) +
+  geom_point(color = "orange") +
+  labs(x = "3PM", y = "Wins", title = "Wins vs 3PM") +
+  theme_minimal() +  # Apply a minimal theme
+  theme(
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.grid.major = element_blank(),         # Remove major gridlines
+    panel.grid.minor = element_blank()          # Remove minor gridlines
+  )
+```
 
-## Discussion
+@fig-2 shows us a visual representation of how 3 point shots made are related to the wins of a team.
+The graph illustrates a positive correlation between the number of 3-point shots made per game and team wins, indicating the importance of proficient long-range shooting in securing victories.
+However, outliers exist, suggesting that factors beyond 3-point shooting also influence team success.
+We discuss these in detail in the discussion portion.
+
+If we group our statistics by team (@cite-groupby), we can gain valuable insights into which teams benefit most from their three-point shooting prowess.
+@tbl-3 presents a detailed breakdown of win statistics and three-point shooting statistics grouped by team for the seasons of interest.
+The table is sorted in descending order of wins, allowing us to identify the top-performing teams.
+By examining this table, we can discern trends, such as the correlation between three-point shooting and team success, and identify standout performers in each category.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-3
+#| tbl-cap: "Table showing the win stats and 3 point stats grouped by the Team"
+#| warning: false
+
+a <- grouped_data
+colnames(a) <- c('Team', 'Games', 'Wins','3PM', '3PA', 'Avg Wins', 'Avg 3PM')
+sorted_table <- a %>%
+  arrange(desc(Wins))
+
+# Display the sorted table
+kable(sorted_table)
+```
+
+```{r}
+#| echo: false
+#| label: fig-1
+#| fig-cap: "Graph depicting the **Average team wins** and **predicted team wins** vs **Average three point shots made** per game"
+#| warning: false
+# Calculate predicted wins
+predicted_wins <- fitted(nba_model2)
+
+# Create the plot
+ggplot(grouped_data, aes(x = Average_3PM, y = Average_Wins)) +
+  geom_point(color = "black") +
+  geom_line(aes(y = predicted_wins, color = "Predicted Wins")) +
+  labs(x = "Average 3PM", y = "Average Wins", title = "Average Wins vs. Average 3PM") +
+  theme_minimal() +  # Apply a minimal theme
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    axis.line = element_line(colour = "black"),  # Set color of axis lines
+    text = element_text(size = 12),  # Adjust text size
+    plot.title = element_text(hjust = 0.5),  # Center the plot title
+    legend.position = "top"  # Set legend position
+  ) +
+  scale_color_manual(values = "orange", name = "")  # Set legend title and color
+```
+
+@fig-1, plotted using ggplot (@cite-ggplot) illustrates the association between the average number of 3-point shots made per game and the average number of wins per season for NBA teams.
+The black points represent observed average wins, while the orange line depicts predicted wins based on a statistical model.
+This model suggests a relationship between 3-point shooting proficiency and team success.
+Comparing the orange line with the black points allows for an evaluation of the model's predictive accuracy.
+
+# Discussion
+
+In this study, we aimed to investigate the relationship between three-point shooting efficiency and team success in the NBA.
+Our analysis involved a comprehensive examination of season-wise NBA statistics, focusing on variables such as three-pointers made (3PM) and team wins (W).
+We employed advanced statistical techniques, including data cleaning, correlation analysis, and regression modeling, to uncover insights into the impact of three-point shooting on team performance.
+By analyzing the data and modeling the relationship between 3PM and W, we sought to shed light on the significance of three-point shooting as a predictor of NBA team success.
+
+Our analysis revealed a compelling association between three-point shooting proficiency and team success in the NBA.
+We found that teams with higher rates of three-pointers made per game tended to achieve more wins over the course of the regular season.
+This suggests that three-point shooting plays a crucial role in determining the outcome of NBA games, with proficient long-range shooters contributing significantly to their team's overall success.
+These findings underscore the strategic importance of three-point shooting in modern basketball and highlight its impact on team performance.
+Additionally, our analysis uncovered nuanced insights into the dynamics of three-point shooting and team success.
+We observed variations in the relationship between 3PM and W across different teams and seasons, indicating that the impact of three-point shooting may vary based on contextual factors such as team composition, playing style, and opponent strategies.
+Furthermore, our findings suggest that while three-point shooting is a significant predictor of team success, it is not the sole determinant, with other factors such as defense, rebounding, and overall offensive efficiency also playing crucial roles in determining game outcomes.
+
+## Weaknesses
+
+Despite the rigor of our analysis, several limitations warrant consideration.
+One potential limitation is the reliance on aggregate team statistics, which may overlook individual player contributions and team dynamics that could influence the relationship between 3PM and W.
+Additionally, our analysis focused solely on regular-season data, excluding playoff games, which may have different dynamics and performance factors since the game is known to "slow down" in the playoffs.
+This can be caused by teams specifically executing strategies to identify and take advantage of the opponents weaknesses over a 7 game series.
+
+Furthermore, while our modeling approach accounted for various factors, including team performance metrics and game outcomes, it may not capture all relevant variables that could impact the relationship between three-point shooting and team success.
+Success in the NBA cannot simply be boiled down to efficient volume three point shooting since it is a nuanced game with many other factors contributing to a team's overall success.
+
+## Drawbacks of the 3 point revolution
 
@@
 An interesting example of this in the modern NBA is the 2017-18 Houston Rockets.
-The Houston Rockets' general manager was a huge fan of the three point shot and in the 2017-18 season built an entire roster around 3 point shooting.
+The Houston Rockets' general manager was a huge fan of the three point shot and in the 2017-18 season built an entire roster around 3 point shooting (@cite-Rockets1).
 This team was a huge success in the regular season and they shot more 3 pointers than any other team in league history.
 What he failed to take into account was, that basketball isn't math.
-While the Rockets were the most successful regular season team with 65 wins, they matched up against the Golden State Warriors in the western conference finals of the playoffs.
+While the Rockets were the most successful regular season team with 65 wins (@cite-Rockets2), they matched up against the Golden State Warriors in the western conference finals of the playoffs.
 The series went to seven games, and in game 7, the Rockets essentially shot themselves out of the game.
@@
 This team lived by the three and eventually would die by the three.
-This just goes to show that although the three point shot is a great indicator of success, it is not without it's drawbacks.
 
-## References
+In essence, while the three-point shot remains a potent weapon in the modern NBA, its effectiveness is contingent upon a nuanced understanding of its role within a team's overall offensive strategy.
+As the Houston Rockets learned, basketball is not solely a game of statistics—factors such as matchup dynamics, defensive schemes, and clutch performances can influence outcomes in ways that transcend numerical analysis.
+
+## Future Research
+
+Looking ahead, there are plenty of opportunities for future research to build upon our discoveries and tackle the limitations of our study.
+One promising avenue is to delve deeper into the data by analyzing individual player-level statistics.
+This approach could help us better grasp how specific players contribute to their team's success, providing a more nuanced understanding of the role of each player in three-point shooting dynamics.
+
+Moreover, conducting longitudinal studies that span multiple seasons could offer valuable insights into the evolving trends and patterns of three-point shooting in the NBA.
+By tracking team performance over time, researchers could uncover long-term dynamics and factors that influence the effectiveness of three-point strategies.
+
+Additionally, qualitative research methods, such as interviews with players and coaches, present an opportunity to gain unique perspectives on the strategic significance of three-point shooting.
+Engaging directly with those involved in the game could offer valuable insights into the decision-making processes and team dynamics surrounding three-point strategies.
+
+By exploring these avenues of inquiry, future research has the potential to deepen our understanding of the intricate relationship between three-point shooting and team success in the NBA, ultimately contributing to advancements in basketball analytics and strategy.
+
+# References

### 2024-04-12 scripts/02-data_cleaning.R

@@ -7,6 +7,7 @@
 # Pre-requisites: none
 
 library(dplyr)
+library(readxl)
 library(openxlsx)
 # Read raw data
 raw_data <- read_excel("data/raw_data/raw_data2.xlsx")
@@ -14,5 +15,17 @@ raw_data <- read_excel("data/raw_data/raw_data2.xlsx")
 # Filter data to only include wins and 3 point shooting stats
 analysis_data <- raw_data[, c("TEAM", "GP", "W", "L", "WIN%", "3PM", "3PA", "3P%", "+/-")]
 
+# Group the stats by team for all the seasons
+grouped_data <- analysis_data %>%
+  group_by(TEAM) %>%
+  summarize(
+    Total_GP = sum(GP),
+    Total_Wins = sum(W),
+    Total_3PM = sum(`3PM`),
+    Total_3PA = sum(`3PA`),
+    Average_Wins = mean(W),
+    Average_3PM = mean(`3PM`)
+  )
 # Write cleaned data to CSV
 write.csv(analysis_data, "data/analysis_data/analysis_data.csv", row.names = FALSE)
+write.csv(grouped_data, "data/analysis_data/grouped_data.csv", row.names = FALSE)

### 2024-04-12 scripts/03-model.R

@@ -12,6 +12,7 @@ library(readr)
 
 #### Read data ####
 analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+grouped_data <- read_csv("data/analysis_data/grouped_data.csv")
 
 ### Model data ####
 # Assuming 'analysis_data' contains columns: 'Season', 'TEAM', 'GP', 'W', 'L', '3PM', '3PA', '3P%'
@@ -28,8 +29,25 @@ nba_model <- stan_glm(
   seed = 853
 )
 
-#### Save model ####
+# Cumulative model based on each Team
+# Season-wise model
+nba_model2 <- stan_glm(
+  formula = Average_Wins ~ Average_3PM,
+  data = grouped_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
+
+#### Save models ####
 saveRDS(
   nba_model,
   file = "model/nba_model.rds"
+)
+
+saveRDS(
+  nba_model2,
+  file = "model/nba_model_team.rds"
 )

### 2024-04-13 paper/paper.qmd
Update paper.qmd

@@
 This study quantifies this link using Generalized Linear Models, which take into account parameters such as the amount of three-point shots made and other pertinent variables.
-The findings shed light on the importance of three-point shooting as an NBA success factor.
+The findings exemplify the importance of three-point shooting as an NBA success factor.
 I examine the ramifications of the findings for team strategy, player development, and basketball analytics.
@@
 My analysis aims to uncover insights into the factors contributing to team success in the NBA, specifically 3 point shooting.
-The NBA's database serves as a comprehensive repository of statistical information on teams, players, and games, providing valuable insights into the dynamics of professional basketball.
+The NBA's database serves as a complete repository of statistical information on teams, players, and games, providing valuable insights into the dynamics of professional basketball.
 Below, I provide a detailed overview of the data collection process, cleaning procedures, and subsequent analysis.
@@
 
-All raw data utilized in this study was collected from the official NBA database (@cite-NBA), which aggregates comprehensive statistics on team performance, player performances, and game outcomes throughout the league's history.
+All raw data utilized in this study was collected from the official NBA database (@cite-NBA), which aggregates all statistics on team performance, player performances, and game outcomes throughout the league's history.
 Since this study only focuses on the modern NBA, which is defined as the era from the 2014-15 season onward, the season wise data had to be individually downloaded and combined.
@@
 By prioritizing **3PM** alongside Wins, our model aims to uncover nuanced insights into the impact of three-point scoring on overall team success while mitigating potential biases associated with alternative metrics.
-This strategic selection ensures a comprehensive exploration of the relationship between three-point performance and winning outcomes in NBA games.
+This strategic selection ensures a throrough exploration of the relationship between three-point performance and winning outcomes in NBA games.
 
@@
 ```
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-3
+#| fig-cap: "Graph showing the average 3 point shots made per game across all the seasons"
+#| warning: false
 
+team_colors <- c(
+  "Atlanta Hawks" = "#E03A3E",
+  "Boston Celtics" = "#007A33",
+  "Brooklyn Nets" = "#000000",
+  "Charlotte Hornets" = "#00788C",
+  "Chicago Bulls" = "#CE1141",
+  "Cleveland Cavaliers" = "#860038",
+  "Dallas Mavericks" = "#00538C",
+  "Denver Nuggets" = "#0E2240",
+  "Detroit Pistons" = "#C8102E",
+  "Golden State Warriors" = "#1D428A",
+  "Houston Rockets" = "#CE1141",
+  "Indiana Pacers" = "#002D62",
+  "LA Clippers" = "#C8102E",
+  "Los Angeles Lakers" = "#552583",
+  "Memphis Grizzlies" = "#5D76A9",
+  "Miami Heat" = "#98002E",
+  "Milwaukee Bucks" = "#00471B",
+  "Minnesota Timberwolves" = "#0C2340",
+  "New Orleans Pelicans" = "#0C2340",
+  "New York Knicks" = "#006BB6",
+  "Oklahoma City Thunder" = "#007AC1",
+  "Orlando Magic" = "#0077C0",
+  "Philadelphia 76ers" = "#006BB6",
+  "Phoenix Suns" = "#1D1160",
+  "Portland Trail Blazers" = "#E03A3E",
+  "Sacramento Kings" = "#5A2D81",
+  "San Antonio Spurs" = "#C4CED4",
+  "Toronto Raptors" = "#CE1141",
+  "Utah Jazz" = "#002B5C",
+  "Washington Wizards" = "#002B5C"
+)
+ggplot(grouped_data, aes(x = Average_3PM, y = reorder(TEAM, Average_3PM), fill = TEAM)) +
+  geom_bar(stat = "identity") +
+  geom_text(aes(label = round(Average_3PM, 1)), hjust = 1.2, size = 3, color = "black") +
+  labs(x = "Average 3-Pointers Made", y = "Team") +
+  scale_fill_manual(values = team_colors) +
+  theme(legend.position = "none")
+```
+@fig-3 shows the average 3 point shots made per game across the 8 seasons that are involved in our study.
 ```{r}
@@
 In this study, we aimed to investigate the relationship between three-point shooting efficiency and team success in the NBA.
-Our analysis involved a comprehensive examination of season-wise NBA statistics, focusing on variables such as three-pointers made (3PM) and team wins (W).
+Our analysis involved an thorough examination of season-wise NBA statistics, focusing on variables such as three-pointers made (3PM) and team wins (W).
 We employed advanced statistical techniques, including data cleaning, correlation analysis, and regression modeling, to uncover insights into the impact of three-point shooting on team performance.
-By analyzing the data and modeling the relationship between 3PM and W, we sought to shed light on the significance of three-point shooting as a predictor of NBA team success.
+By analyzing the data and modeling the relationship between 3PM and W, we sought to show the significance of three-point shooting as a predictor of NBA team success.
 
@@
 We found that teams with higher rates of three-pointers made per game tended to achieve more wins over the course of the regular season.
-This suggests that three-point shooting plays a crucial role in determining the outcome of NBA games, with proficient long-range shooters contributing significantly to their team's overall success.
+This suggests that three-point shooting plays a key role in determining the outcome of NBA games, with proficient long-range shooters contributing significantly to their team's overall success.
 These findings underscore the strategic importance of three-point shooting in modern basketball and highlight its impact on team performance.
@@
 We observed variations in the relationship between 3PM and W across different teams and seasons, indicating that the impact of three-point shooting may vary based on contextual factors such as team composition, playing style, and opponent strategies.
-Furthermore, our findings suggest that while three-point shooting is a significant predictor of team success, it is not the sole determinant, with other factors such as defense, rebounding, and overall offensive efficiency also playing crucial roles in determining game outcomes.
+Furthermore, our findings suggest that while three-point shooting is a significant predictor of team success, it is not the sole determinant, with other factors such as defense, rebounding, and overall offensive efficiency also playing pivotal roles in determining game outcomes.
 
@@
 Looking ahead, there are plenty of opportunities for future research to build upon our discoveries and tackle the limitations of our study.
-One promising avenue is to delve deeper into the data by analyzing individual player-level statistics.
+One promising avenue is to look deeper into the data by analyzing individual player-level statistics.
 This approach could help us better grasp how specific players contribute to their team's success, providing a more nuanced understanding of the role of each player in three-point shooting dynamics.
@@
 
-By exploring these avenues of inquiry, future research has the potential to deepen our understanding of the intricate relationship between three-point shooting and team success in the NBA, ultimately contributing to advancements in basketball analytics and strategy.
+By exploring these avenues of inquiry, future research has the potential to deepen our understanding of the intricate relationship between three-point shooting and team success in the NBA, contributing to advancements in basketball analytics and strategy.
 

### 2024-04-17 other/datasheet/datasheet.qmd
Adding the datasheet

---
title: "Datasheet for nba.com regular season team stats"
subtitle: "The dataset includes regular season data of the traditional stats from the 2014-15 season to the 2022-23 season (excluding the 2019-20 season)"
author: 
  - Aryaman Suri
thanks: "Code and data are available at: https://github.com/aryamansuri/NBA-3PointAnalysis"
date: today
date-format: long
abstract: "The NBA Traditional Team Statistics dataset offers a comprehensive collection of data covering the performance of NBA teams from the 2014-15 season to the 2022-23 season, excluding 2019-20. Curated from NBA.com (@cite-NBA), it includes various traditional team statistics metrics, providing valuable insights into team performance dynamics over time. This dataset serves as a crucial resource for sports analytics, research, and exploratory data analysis in the field of basketball statistics. Its availability on NBA.com, the official website of the National Basketball Association, underscores its reliability and relevance for studying trends and patterns in professional basketball."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The NBA Traditional Team Statistics dataset was compiled to provide comprehensive insights into the performance of NBA teams during the regular season. Its creation aimed to offer a detailed overview of various traditional team statistics metrics, including total games played, wins, 3-pointers made, and other key performance indicators. By aggregating this data from NBA.com, the dataset facilitates detailed analysis of team dynamics, trends, and patterns over multiple seasons, enabling researchers, analysts, and basketball enthusiasts to gain deeper insights into team performance and strategies.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was sourced from NBA.com, the official website of the National Basketball Association (NBA), which is responsible for compiling and publishing comprehensive statistics and information related to NBA games. Therefore, the dataset can be attributed to the NBA organization itself, as it represents the official records and data collected by the league.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of the dataset was funded by the National Basketball Association (NBA) as part of its ongoing operations and initiatives to collect and maintain comprehensive statistics for analysis and reporting purposes. As such, there may not be a specific grant associated with the creation of this dataset, as it falls under the regular activities and operations of the NBA.
4. *Any other comments?*
    - No

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances in the dataset represent NBA teams and their corresponding statistics across multiple seasons, specifically focusing on traditional team statistics such as games played, wins, field goals made, three-pointers made, rebounds, assists, steals, blocks, turnovers, and more. Each instance corresponds to a single NBA team in a particular season, providing a comprehensive overview of their performance metrics during the regular season.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The total number of instances in the dataset depends on the number of NBA teams and the number of seasons included in the dataset. Typically, there are 30 NBA teams in each season, resulting in a variable number of instances depending on the range of seasons covered. For example, if the dataset covers eight seasons (excluding the 2019-20 season), there would be a total of 240 instances (30 teams * 8 seasons).
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
  - The dataset comprises a comprehensive sample of NBA team statistics extracted from the NBA.com traditional team stats for the regular seasons spanning from 2014-15 to 2022-23, excluding the 2019-20 season. While it does not encompass all possible instances within this timeframe, it offers a representative snapshot of team performances across multiple seasons. The sample's representativeness is ensured by its inclusion of data from various teams and seasons, providing a broad overview of NBA team statistics during the specified period. However, certain factors, such as variations in team performance over time and potential biases in data collection methods, may affect its complete representativeness. Nonetheless, efforts were made to include diverse teams and seasons to mitigate potential biases and enhance the dataset's overall utility for analysis and research.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
  - Each instance in the dataset consists of aggregated statistical features representing the performance of individual NBA teams in various categories throughout the regular seasons from 2014-15 to 2022-23, excluding the 2019-20 season. These features include metrics such as total games played, total wins, total 3-pointers made (3PM), total 3-point attempts (3PA), average wins, and average 3-pointers made (Average_3PM). The data is structured in a tabular format, with each row representing a specific team in a particular season, and each column containing a different statistical attribute or feature.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, there is a target associated with each instance in the dataset. The target variable is the average number of 3-pointers made (Average_3PM) by each NBA team during the regular season. This serves as a key metric for assessing the performance of teams in terms of their proficiency in making 3-point shots.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
  - No, there is no missing information from individual instances in the dataset. Each instance includes complete data for the variables relevant to the analysis, such as the team name (TEAM), total games played (Total_GP), total wins (Total_Wins), total 3-pointers made (Total_3PM), total 3-point attempts (Total_3PA), average wins (Average_Wins), and average 3-pointers made (Average_3PM). There are no instances with missing or unavailable information.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- No explicit relationships between individual instances are present in the dataset. Each instance represents aggregated statistical information for a specific NBA team in a given season.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- There are no recommended data splits provided with the dataset.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No specific errors, sources of noise, or redundancies have been identified in the dataset. However, typical data quality checks and preprocessing steps should be performed before analysis to ensure accuracy.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset does not rely on external resources. All data included in the dataset are self-contained and sourced from NBA.com.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, the dataset does not contain data that would be considered confidential. It consists solely of aggregated statistical information related to NBA team performance, which is publicly available data.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No, the dataset does not contain any content that might be offensive, insulting, threatening, or cause anxiety. It consists entirely of statistical data related to NBA team performance.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The dataset does not identify specific sub-populations such as age or gender. Each instance represents aggregated statistics for individual NBA teams across different seasons.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No, it is not possible to identify individuals directly or indirectly from the dataset. The dataset consists of aggregated team-level statistics and does not contain any personal identifiers or individual-level data.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No, the dataset does not contain any sensitive data. It consists solely of statistical information related to NBA team performance and does not include any personal, sensitive, or demographic information.
16. *Any other comments?*
	- No

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data associated with each instance was acquired from NBA.com's traditional team statistics for the regular seasons spanning from the 2014-15 season to the 2022-23 season, excluding the 2019-20 season. This data is directly observable and publicly available, comprising aggregated statistical information for each NBA team in a given season.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The data collection process involved accessing and extracting statistical information from NBA.com's traditional team statistics pages for each season. This process was conducted manually by accessing the relevant web pages and recording the required statistics. The accuracy of the data was ensured through careful verification and cross-referencing with multiple sources to mitigate potential errors or inconsistencies.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset represents a comprehensive collection of NBA team statistics for the specified seasons, excluding the 2019-20 season. It is not a sample but rather encompasses data for all NBA teams across multiple seasons.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The data collection process was carried out by individuals manually accessing and recording statistical information from NBA.com. There were no specific compensations involved, as the data collection was conducted as part of research or analytical activities.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected over multiple seasons, spanning from the 2014-15 season to the 2022-23 season, excluding the 2019-20 season. The creation timeframe of the data associated with the instances aligns with the respective NBA regular seasons in which the statistics were recorded.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No specific ethical review processes were conducted for the acquisition of publicly available NBA team statistics. The data collection process adhered to ethical standards and guidelines applicable to web scraping and data analysis activities.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was obtained directly from NBA.com, a reputable source for NBA-related statistics and information. No intermediary parties were involved in the data collection process.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- As the data was collected from a publicly accessible website (NBA.com), no explicit notification to individuals was required or provided. NBA.com's terms of service and privacy policy govern the collection and use of data from their platform.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Since the data was publicly available on NBA.com and accessible to all users, explicit consent from individuals was not required for its collection and use. NBA.com's terms of service and privacy policy govern the use of data available on their platform.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- N/A
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- N/A
12. *Any other comments?*
	- No

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- The dataset underwent preprocessing to ensure consistency and accuracy. This included standardizing team names, handling missing values, and aggregating statistics for each team across multiple seasons. 
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, the raw data obtained directly from NBA.com's traditional team statistics pages was saved alongside the preprocessed and cleaned dataset. The raw data serves as a reference point and can be accessed upon request for transparency and reproducibility purposes.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The software used for preprocessing and cleaning the data is not publicly available. However, the preprocessing steps were implemented using commonly used data manipulation and cleaning tools in the R programming language, such as dplyr and tidyr packages. Detailed documentation of the preprocessing steps is available upon request for transparency and reproducibility.
4. *Any other comments?*
	- No

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset has been utilized in various analytical tasks within the domain of sports analytics. It has been employed to analyze the relationship between 3-point field goals made and team success, identify trends in shooting performance across different NBA seasons, and assess the impact of key performance indicators on game outcomes. Additionally, researchers have explored the dataset to understand the evolution of playing styles and strategies employed by NBA teams over time.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- As of now, there is no centralized repository that links to papers or systems utilizing the dataset. However, individual research papers and analytical reports utilizing the dataset may be found through academic databases, sports analytics journals, or online repositories.
3. *What (other) tasks could the dataset be used for?*
	- The dataset presents opportunities for various tasks beyond sports analytics. It could be leveraged for predictive modeling to forecast team performance in upcoming seasons, player performance evaluation, and talent scouting. Furthermore, the dataset could serve as a case study for statistical analysis courses, providing real-world examples of data analysis and visualization techniques.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- While the dataset primarily consists of aggregated team statistics, users should exercise caution in drawing conclusions about individual players or making generalized statements about team performance based solely on statistical aggregates. Additionally, it's important to consider the contextual factors surrounding each NBA season, such as rule changes, team dynamics, and external influences, which may impact the interpretation of the data. Consumers should approach the dataset with an understanding of its limitations and avoid making overly deterministic assertions.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- While the dataset provides valuable insights into team performance and statistical trends within the NBA, it should not be used as the sole basis for making high-stakes decisions, such as betting on games or player evaluations for professional scouting purposes. Users should avoid extrapolating the findings of the dataset beyond the scope of sports analytics without considering additional contextual information and expert knowledge.
6. *Any other comments?*
	- No

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The distribution of the dataset to third parties outside the entity is currently under consideration. If approved, distribution will be conducted following established protocols to ensure proper usage and adherence to any applicable terms of use.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset's distribution method is yet to be finalized. Potential distribution channels include hosting on a website, publication on platforms like GitHub, or through APIs. Efforts will be made to assign a digital object identifier (DOI) to facilitate citation and proper attribution.
3. *When will the dataset be distributed?*
	- The timeline for dataset distribution is currently being determined and will be communicated once finalized.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- The dataset may be distributed under a specific copyright or intellectual property (IP) license, along with applicable terms of use (ToU). Details regarding the license, ToU, and any associated fees will be provided upon distribution.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- Any IP-based or other restrictions imposed by third parties on the dataset will be communicated transparently to dataset users. Relevant licensing terms and any associated fees will be provided along with the dataset.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- Export controls or other regulatory restrictions, if applicable, will be clearly outlined to ensure compliance with relevant laws and regulations. Documentation supporting these restrictions will be made available to dataset consumers.
7. *Any other comments?*
	- No

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The entity responsible for the creation of the dataset, which is the NBA, will also oversee its hosting, maintenance, and support.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- The NBA can be contacted via a contact form via https://contact.nba.com/contact-nba/
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Updates to the dataset, including corrections, additions, or deletions, will be managed by the dataset's custodian. The frequency of updates and the process for communicating these changes to dataset consumers will be determined and clearly communicated.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- Any applicable limits on the retention of personal data associated with the dataset will be communicated in accordance with relevant privacy laws and regulations. Procedures for enforcing these limits will be implemented to ensure compliance.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Plans for the support and maintenance of older versions of the dataset are under consideration. If supported, access to older versions will be facilitated through archival mechanisms or version control systems. In the event of obsolescence, clear communication channels will be established to notify dataset consumers.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Mechanisms for extending, augmenting, or contributing to the dataset may be established, subject to approval and validation processes. Contributors will be required to adhere to specific guidelines and standards to ensure the integrity and quality of contributions. A transparent process for communicating and distributing these contributions to dataset consumers will be implemented to facilitate collaboration and knowledge sharing.
8. *Any other comments?*
	- No

\newpage

# References

### 2024-04-17 paper/paper/paper.qmd
Add files via upload

--- 2024-04-13 paper/paper.qmd
@@
 
-The three point line was introduced in the NBA in the 1979-80 season and has brought unprecedented balance to the game.
+The three point line was introduced in the NBA in the 1979-80 season (@cite-threepoint) and has brought unprecedented balance to the game.
 Instead of the game being centered around the big man or athletic wings, the 3 point shot gave everybody on the team a new powerful weapon on offenses.
-Today, there are 230 players who averaged at least one made three pointer in the 2022-23 regular season.
+Today, there are 155 players who averaged at least one made three pointer in the 2022-23 regular season (@cite-threepoint2).
 The estimand of this paper is to see if these 3 point shots help the team win games in the regular season.
 
-This paper aims to estimate the association between three-point shooting measures and win probability using advanced statistical techniques after meticulously analyzing season-wise data.
-The findings highlighted the importance of three-point shooting efficiency as a predictor of team performance in modern basketball and how important it is to be a good three point shooting team for success in today's game.
+This paper explores the association between three-point shooting measures and win probability using advanced statistical techniques after meticulously analyzing season-wise data.
+The findings highlight the importance of three-point shooting efficiency as a predictor of team performance in modern basketball and how important it is to be a good three point shooting team for success in today's game.
 
@@
 This study adds to the ongoing debate in the NBA about strategic decision-making and player development by explaining the significance of three-point shooting in deciding game outcomes.
-Essentially, the question "Does three point shooting actually lead to team success?", is being asked.
+Essentially, the question “Does three point shooting actually lead to team success?”, is being asked.
 
-I begin by exploring the underlying data, which will provide insights into the variables and metrics used in our analyses.
-Following that, I explain the complexities of the modelling approach, describing the reasoning behind the statistical techniques and justifying the methodology.
-Later, the results of the analysis are discussed, diving into the empirical findings and their implications for understanding the relationship between three-point shooting efficiency and NBA team success.
-Then  I have included a thorough conversation, contextualizing the findings within the larger landscape of basketball analytics and emphasizing significant takeaways from this research.
-Finally, I summarize the findings and propose paths for future research to further the understanding of this dynamic interaction.
+This paper begins by exploring the underlying data in @sec-data, which provides insights into the variables and metrics used in the analyses.
+Following that, the complexities of the modelling approach are explained in @sec-model, including describing the reasoning behind the statistical techniques and justifying the methodology.
+The results of the analysis are discussed in @sec-results, diving into the empirical findings and their implications for understanding the relationship between three-point shooting eﬀiciency and NBA team success.
+@sec-discuss contextualizes the findings of this paper within the larger landscape of basketball analytics,emphasizing significant takeaways.
+Finally, the findings and proposed paths for future research are summarized in @sec-discuss2.
 
-# Data
+# Data {#sec-data}
 
-In this study, I utilized data sourced from the NBA's official database (@cite-NBA) to examine the relationship between various team performance metrics and their corresponding winning percentages throughout the regular season.
-My analysis aims to uncover insights into the factors contributing to team success in the NBA, specifically 3 point shooting.
-The NBA's database serves as a complete repository of statistical information on teams, players, and games, providing valuable insights into the dynamics of professional basketball.
-Below, I provide a detailed overview of the data collection process, cleaning procedures, and subsequent analysis.
+This study, utilizes  data sourced from the NBA’s oﬀicial database (@cite-NBA) to examine the relationship between various team performance metrics and their corresponding winning percentages throughout the regular season.
+The analysis aims to uncover insights into the factors contributing to team success in the NBA, specifically 3 point shooting.
+The NBA’s database serves as a complete repository of statistical information on teams, players, and games, providing valuable insights into the dynamics of professional basketball.
+The below sections provide  a detailed overview of the data collection process, cleaning procedures, and subsequent analysis.
 
@@
 
-All raw data utilized in this study was collected from the official NBA database (@cite-NBA), which aggregates all statistics on team performance, player performances, and game outcomes throughout the league's history.
-Since this study only focuses on the modern NBA, which is defined as the era from the 2014-15 season onward, the season wise data had to be individually downloaded and combined.
-Additionally, since the 2019-20 season was shortened due to the COVID-19 outbreak and an equal number of games was not played by all teams, we will not be considering that season for the sake of this study.
-This data is then stored in our raw_data2.xlsx file using @cite-writexl.
-This dataset comprises a wide array of variables, including team statistics such as points scored, rebounds, assists, turnovers, and various other performance metrics recorded over multiple seasons.
+All raw data utilized in this study was collected from the oﬀicial NBA database  (@cite-NBA), which aggregates statistics on team performance, player performances, and game outcomes throughout the league’s history.
+Since this study only focuses on the modern NBA, which is defined as the era from the 2014-15 season onward (@cite-modernera), the season wise data had to be individually downloaded and combined.
+Additionally, since the 2019-20 season was shortened due to the COVID-19 outbreak and an equal number of games was not played by all teams (@cite-shortseason), data from that season is excluded from this study to ensure consistency.
+The aggregated  dataset is then stored as the raw_data2.xlsx file using the *writexl* package (@cite-writexl).
+This dataset comprises data from 30 teams over 8 seasons, including team statistics such as points scored, rebounds, assists, turnovers, and various other performance metrics recorded over multiple seasons.
 
@@
 
-Upon acquiring the raw data from nba.com (@cite-NBA), extensive cleaning procedures were implemented to ensure its suitability for analysis.
+Upon acquiring the raw data from NBA.com(@cite-NBA), extensive cleaning procedures were implemented to ensure its suitability for analysis.
 This involved addressing missing or erroneous values, standardizing variable names, and aggregating data at the team level to facilitate meaningful comparisons across different seasons and teams.
@@
 
-# Model
+# Model {#sec-model}
 
@@
 This approach allows for a nuanced understanding of how variations in 3-point performance contribute to variations in team outcomes, thereby elucidating the importance of 3-point shooting in the context of NBA team success.
+Simply put, teams that excel in scoring 3-point shots are more likely to outscore their opponents and win the games.
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-discuss}
 
@@
 
-## Future Research
+## Future Research {#sec-discuss2}
 
@@
 
+{{< pagebreak >}}
+
 # References

### 2024-04-18 paper/paper.qmd
Final changes

@@
 ---
-title: "Beyond the Arc: How 3 point shooting has impacted winning in the modern NBA"
-subtitle: "Analysing the relationship between a team's 3 point shooting and sucess in the regular season"
+title: "Beyond the Arc: The Impact of  3 point Shooting in the Modern NBA"
+subtitle: "Analysing the relationship between a team’s 3 point shooting and their success in the regular season"
 author: 
@@
 date-format: long
-abstract: "In recent years, the three-point shot has emerged as a pivotal aspect of basketball strategy in the NBA.
-This study uses statistical modelling tools to analyze the association between three-point shooting and winning in the current NBA.
-Using season-wise NBA statistics, I examine the cumulative effect of teams' three-point shooting on their overall victory total.
-This study quantifies this link using Generalized Linear Models, which take into account parameters such as the amount of three-point shots made and other pertinent variables.
-The findings exemplify the importance of three-point shooting as an NBA success factor.
-I examine the ramifications of the findings for team strategy, player development, and basketball analytics.
-Additionally, ethical concerns and study limitations are addressed."
+abstract: "In recent years, the three-point shot has emerged as a key aspect of basketball strategy in the NBA.
+Analyzing season-wise NBA data, we assess how teams' three-point performance influences their overall wins.
+Employing Bayesian Generalized Linear Model, we quantify this relationship, considering factors like team three-point averages across regular seasons.
+The findings underscore the significance of three-point shooting in NBA success, offering insights into team strategies, player growth, and basketball analytics.
+Ethical considerations and study constraints are also addressed."
 format: pdf
@@
 
-The three point line was introduced in the NBA in the 1979-80 season and has brought unprecedented balance to the game.
+The three point line was introduced in the NBA in the 1979-80 season (@cite-threepoint) and has since altered the nature and balance of the game.
 Instead of the game being centered around the big man or athletic wings, the 3 point shot gave everybody on the team a new powerful weapon on offenses.
-Today, there are 230 players who averaged at least one made three pointer in the 2022-23 regular season.
-The estimand of this paper is to see if these 3 point shots help the team win games in the regular season.
+Today, there are 155 players who averaged at least one made three pointer in the 2022-23 regular season (@cite-threepoint2).
 
-This paper aims to estimate the association between three-point shooting measures and win probability using advanced statistical techniques after meticulously analyzing season-wise data.
-The findings highlighted the importance of three-point shooting efficiency as a predictor of team performance in modern basketball and how important it is to be a good three point shooting team for success in today's game.
+This paper explores the association between three-point shooting measures and win probability using advanced statistical techniques after analyzing season-wise data.
+The findings highlight the importance of three-point shooting eﬀiciency as a predictor of team performance in modern basketball and how important it is to be a good three point shooting team for success in today’s game.
 
-Understanding this link has far-reaching ramifications for numerous parties, including coaches, players, analysts, and team administration.
-This study adds to the ongoing debate in the NBA about strategic decision-making and player development by explaining the significance of three-point shooting in deciding game outcomes.
-Essentially, the question "Does three point shooting actually lead to team success?", is being asked.
+Understanding this link has ramifications for numerous parties, including coaches, players, analysts, and team administration.
+This study adds to the ongoing debate in the NBA about strategic decision-making and player development by explaining the significance of three-point shooting in deciding game outcomes (@cite-debate).
+Essentially, the question “Does three point shooting actually lead to team success?”, is being asked.
 
-I begin by exploring the underlying data, which will provide insights into the variables and metrics used in our analyses.
-Following that, I explain the complexities of the modelling approach, describing the reasoning behind the statistical techniques and justifying the methodology.
-Later, the results of the analysis are discussed, diving into the empirical findings and their implications for understanding the relationship between three-point shooting efficiency and NBA team success.
-Then  I have included a thorough conversation, contextualizing the findings within the larger landscape of basketball analytics and emphasizing significant takeaways from this research.
-Finally, I summarize the findings and propose paths for future research to further the understanding of this dynamic interaction.
+This paper begins by exploring the underlying data in @sec-data, which provides insights into the variables and metrics used in the analyses.
+Following that, the complexities of the modelling approach are explained in @sec-model, including describing the reasoning behind the statistical techniques and justifying the methodology.
+The results of the analysis are discussed in @sec-results, diving into the empirical findings and their implications for understanding the relationship between three-point shooting eﬀiciency and NBA team success.
+@sec-discuss contextualizes the findings of this paper within the larger view of basketball analytics,emphasizing significant takeaways.
+Finally, the findings and proposed paths for future research are summarized in @sec-discuss2.
 
-# Data
+# Data {#sec-data}
 
-In this study, I utilized data sourced from the NBA's official database (@cite-NBA) to examine the relationship between various team performance metrics and their corresponding winning percentages throughout the regular season.
-My analysis aims to uncover insights into the factors contributing to team success in the NBA, specifically 3 point shooting.
-The NBA's database serves as a complete repository of statistical information on teams, players, and games, providing valuable insights into the dynamics of professional basketball.
-Below, I provide a detailed overview of the data collection process, cleaning procedures, and subsequent analysis.
+This study, utilizes  data sourced from the NBA’s official database (@cite-NBA) to examine the relationship between various team performance metrics and their corresponding winning percentages throughout the regular season.
+The analysis aims to uncover insights into the factors contributing to team success in the NBA, specifically 3 point shooting.
+The NBA’s database serves as a complete repository of statistical information on teams, players, and games, providing valuable insights into the dynamics of professional basketball.
+The below sections provide  a detailed overview of the data collection process, cleaning procedures, and subsequent analysis.
 
@@
 
-While the measurement of this model's data can be criticized by NBA fans , we've deliberately opted for **3PM** over **3PA** and **3P%** for several reasons.
-**3PM** serves as a robust metric, capturing a team's volume of successful three-point shots with reasonable efficiency.
+We've deliberately opted for **3PM** over **3PA** and **3P%** for several reasons.
+**3PM** serves as a accurate metric, capturing a team's volume of successful three-point shots with reasonable efficiency.
 Unlike **3PA**, which might bias the model towards teams overly reliant on three-point attempts irrespective of their success rate, **3PM** provides a balanced representation of both volume and efficiency in three-point shooting.
@@
 
-By prioritizing **3PM** alongside Wins, our model aims to uncover nuanced insights into the impact of three-point scoring on overall team success while mitigating potential biases associated with alternative metrics.
+By prioritizing **3PM** alongside Wins, our model aims to uncover insights into the impact of three-point scoring on overall team success while mitigating potential biases associated with alternative metrics.
 This strategic selection ensures a throrough exploration of the relationship between three-point performance and winning outcomes in NBA games.
 
-# Model
+# Model {#sec-model}
 
@@
 This hypothesis is rooted in the understanding that teams with higher 3-point shooting efficiency typically demonstrate stronger offensive capabilities, which often translate into higher-scoring games and, consequently, more victories.
-This relationship is particularly pertinent in today's NBA landscape, where teams increasingly prioritize offensive strategies and exploit the scoring potential offered by 3-point shots.
+This relationship is particularly pertinent in today's NBA, where teams increasingly prioritize offensive strategies and exploit the scoring potential offered by 3-point shots.
 
@@
 By modeling the relationship between 3-point field goals made and wins using a normal distribution for the response variable, we acknowledge the continuous nature of team wins and seek to capture the variability in win counts across different levels of 3-point shooting proficiency.
-This approach allows for a nuanced understanding of how variations in 3-point performance contribute to variations in team outcomes, thereby elucidating the importance of 3-point shooting in the context of NBA team success.
+This approach allows for an understanding of how variations in 3-point performance contribute to variations in team outcomes, thereby elucidating the importance of 3-point shooting in the context of NBA team success.
+Simply put, teams that excel in scoring 3-point shots are more likely to outscore their opponents and win the games.
 
-# Results
+# Results {#sec-results}
 
@@
 #| label: fig-2
-#| fig-cap: "Graph depicting the **Wins** vs the **3 Points Made** per game"
+#| fig-cap: "Scatterplot depicting the **Wins** vs the **3 Points Made** per game"
 #| warning: false
@@
   geom_bar(stat = "identity") +
-  geom_text(aes(label = round(Average_3PM, 1)), hjust = 1.2, size = 3, color = "black") +
+  geom_text(aes(label = round(Average_3PM, 1)), position = position_nudge(x = 0.5), vjust = 0.5, size = 3, color = "black") +
   labs(x = "Average 3-Pointers Made", y = "Team") +
   scale_fill_manual(values = team_colors) +
-  theme(legend.position = "none")
+  theme(legend.position = "none", axis.text.y = element_text(size = 8))
 ```
@@
 
-# Discussion
+{{< pagebreak >}}
+
+# Discussion {#sec-discuss}
 
 In this study, we aimed to investigate the relationship between three-point shooting efficiency and team success in the NBA.
-Our analysis involved an thorough examination of season-wise NBA statistics, focusing on variables such as three-pointers made (3PM) and team wins (W).
+Our analysis involved a detailed examination of season-wise NBA statistics, focusing on variables such as three-pointers made (3PM) and team wins (W).
 We employed advanced statistical techniques, including data cleaning, correlation analysis, and regression modeling, to uncover insights into the impact of three-point shooting on team performance.
@@
 These findings underscore the strategic importance of three-point shooting in modern basketball and highlight its impact on team performance.
-Additionally, our analysis uncovered nuanced insights into the dynamics of three-point shooting and team success.
+Additionally, our analysis uncovered insights into the dynamics of three-point shooting and team success.
 We observed variations in the relationship between 3PM and W across different teams and seasons, indicating that the impact of three-point shooting may vary based on contextual factors such as team composition, playing style, and opponent strategies.
-Furthermore, our findings suggest that while three-point shooting is a significant predictor of team success, it is not the sole determinant, with other factors such as defense, rebounding, and overall offensive efficiency also playing pivotal roles in determining game outcomes.
+Furthermore, our findings suggest that while three-point shooting is a significant predictor of team success, it is not the sole determinant, with other factors such as defense, rebounding, and overall offensive efficiency also influencing and determining game outcomes.
 
@@
 Furthermore, while our modeling approach accounted for various factors, including team performance metrics and game outcomes, it may not capture all relevant variables that could impact the relationship between three-point shooting and team success.
-Success in the NBA cannot simply be boiled down to efficient volume three point shooting since it is a nuanced game with many other factors contributing to a team's overall success.
+Success in the NBA cannot simply be boiled down to efficient volume three point shooting since it is a complex game with many other factors contributing to a team's overall success.
 
@@
 
-In essence, while the three-point shot remains a potent weapon in the modern NBA, its effectiveness is contingent upon a nuanced understanding of its role within a team's overall offensive strategy.
+In essence, while the three-point shot remains a potent weapon in the modern NBA, its effectiveness is contingent upon an understanding of its role within a team's overall offensive strategy.
 As the Houston Rockets learned, basketball is not solely a game of statistics—factors such as matchup dynamics, defensive schemes, and clutch performances can influence outcomes in ways that transcend numerical analysis.
 
-## Future Research
+## Future Research {#sec-discuss2}
 
@@
 One promising avenue is to look deeper into the data by analyzing individual player-level statistics.
-This approach could help us better grasp how specific players contribute to their team's success, providing a more nuanced understanding of the role of each player in three-point shooting dynamics.
+This approach could help us better grasp how specific players contribute to their team's success, providing a deeper understanding of the role of each player in three-point shooting dynamics.
 
-Moreover, conducting longitudinal studies that span multiple seasons could offer valuable insights into the evolving trends and patterns of three-point shooting in the NBA.
+Moreover, conducting longitudinal studies that span multiple seasons could offer valuable insights into the trends and patterns of three-point shooting in the NBA.
 By tracking team performance over time, researchers could uncover long-term dynamics and factors that influence the effectiveness of three-point strategies.
@@
 
-By exploring these avenues of inquiry, future research has the potential to deepen our understanding of the intricate relationship between three-point shooting and team success in the NBA, contributing to advancements in basketball analytics and strategy.
+By exploring these avenues of inquiry, future research has the potential to deepen our understanding of the relationship between three-point shooting and team success in the NBA, contributing to advancements in basketball analytics and strategy.
+
+{{< pagebreak >}}
 

### 2024-04-18 scripts/03-test_data.R

@@ -1,0 +1,34 @@
+#### Preamble ####
+# Purpose: Tests analysis data
+# Authors: Aryaman Suri
+# Contact: aryaman.suri@mail.utoronto.ca
+# Pre-requisites: -
+
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+
+#### Run the analysis data ####
+analysis_data <- read.csv("data/analysis_data/analysis_data.csv")
+
+
+
+#### Test the data ####
+test_that("Check for the right class", {
+  expect_type(analysis_data$W, "integer")
+  expect_type(analysis_data$X3PM, "double")
+  
+})
+
+
+test_that("Check complete", {
+  expect_true(all(complete.cases(analysis_data)))
+})
+
+
+test_that("Check variable",{
+  expect_true(analysis_data$W |> min() >= 0)
+  expect_true(analysis_data$W |> max() <= 82) # Since there are only 82 games in a season
+})

### 2024-04-18 scripts/04-model.R

@@ -1,0 +1,53 @@
+#### Preamble ####
+# Purpose: Model to predict wins based on season-wise 3-point shots made (3PM) in the NBA.
+# Author: Aryaman Suri
+# Date: 30 March 2024
+# Contact: aryaman.suri@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(readr)
+
+#### Read data ####
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+grouped_data <- read_csv("data/analysis_data/grouped_data.csv")
+
+### Model data ####
+# Assuming 'analysis_data' contains columns: 'Season', 'TEAM', 'GP', 'W', 'L', '3PM', '3PA', '3P%'
+# For this analysis, we'll focus on 'W' (wins) and '3PM' (3-point shots made)
+
+# Season-wise model
+nba_model <- stan_glm(
+  formula = W ~ `3PM`,
+  data = analysis_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
+
+# Cumulative model based on each Team
+# Season-wise model
+nba_model2 <- stan_glm(
+  formula = Average_Wins ~ Average_3PM,
+  data = grouped_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
+
+#### Save models ####
+saveRDS(
+  nba_model,
+  file = "model/nba_model.rds"
+)
+
+saveRDS(
+  nba_model2,
+  file = "model/nba_model_team.rds"
+)

### 2024-04-18 scripts/03-model.R
Delete scripts/03-model.R

@@ -1,53 +1,0 @@
-#### Preamble ####
-# Purpose: Model to predict wins based on season-wise 3-point shots made (3PM) in the NBA.
-# Author: Aryaman Suri
-# Date: 30 March 2024
-# Contact: aryaman.suri@mail.utoronto.ca
-# License: MIT
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-library(readr)
-
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
-grouped_data <- read_csv("data/analysis_data/grouped_data.csv")
-
-### Model data ####
-# Assuming 'analysis_data' contains columns: 'Season', 'TEAM', 'GP', 'W', 'L', '3PM', '3PA', '3P%'
-# For this analysis, we'll focus on 'W' (wins) and '3PM' (3-point shots made)
-
-# Season-wise model
-nba_model <- stan_glm(
-  formula = W ~ `3PM`,
-  data = analysis_data,
-  family = gaussian(),
-  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_aux = exponential(rate = 1, autoscale = TRUE),
-  seed = 853
-)
-
-# Cumulative model based on each Team
-# Season-wise model
-nba_model2 <- stan_glm(
-  formula = Average_Wins ~ Average_3PM,
-  data = grouped_data,
-  family = gaussian(),
-  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_aux = exponential(rate = 1, autoscale = TRUE),
-  seed = 853
-)
-
-#### Save models ####
-saveRDS(
-  nba_model,
-  file = "model/nba_model.rds"
-)
-
-saveRDS(
-  nba_model2,
-  file = "model/nba_model_team.rds"
-)

### 2024-04-18 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -23,8 +23,8 @@ grouped_data <- analysis_data %>%
     Total_Wins = sum(W),
     Total_3PM = sum(`3PM`),
     Total_3PA = sum(`3PA`),
-    Average_Wins = mean(W),
-    Average_3PM = mean(`3PM`)
+    Average_Wins = round(mean(W), 2),
+    Average_3PM = round(mean(`3PM`), 2)
   )
 # Write cleaned data to CSV
 write.csv(analysis_data, "data/analysis_data/analysis_data.csv", row.names = FALSE)

### 2024-04-18 paper/paper.qmd
Update paper.qmd

@@
 This study adds to the ongoing debate in the NBA about strategic decision-making and player development by explaining the significance of three-point shooting in deciding game outcomes (@cite-debate).
-Essentially, the question “Does three point shooting actually lead to team success?”, is being asked.
+Essentially, the estmand of the paper is to see if an increase in a team's three point shooting ability contributes to an increase in the team's success, measured by wins.
 
@@
 
-## Data Collection
+## Data Sources
 
@@
 #| label: tbl-1
-#| tbl-cap: "Overview of the cleaned dataset"
+#| tbl-cap: "Summary table of the cleaned dataset"
 #| warning: false
@@
 In this study, our focus is solely on win counts and three-point statistics.
-With this objective in mind, our key variables of interest include Games Played (**GP**), Wins (**W**), Losses (**L**), Three-Pointers Made per game (**3PM**), Three-Pointers Attempted per game (3PA), and Three-Point Shooting Percentage (**3P%**).
+With this objective in mind, our key variables of interest include Games Played (**GP**), Wins (**W**), Losses (**L**), Three-Pointers Made per game (**3PM**), Three-Pointers Attempted per game (**3PA**), and Three-Point Shooting Percentage (**3P%**).
 Our modeling approach will specifically incorporate Wins (**W**) and **3PM** to explore their relationship.
@@
 By prioritizing **3PM** alongside Wins, our model aims to uncover insights into the impact of three-point scoring on overall team success while mitigating potential biases associated with alternative metrics.
-This strategic selection ensures a throrough exploration of the relationship between three-point performance and winning outcomes in NBA games.
+This strategic selection ensures a detailed exploration of the relationship between three-point performance and winning outcomes in NBA games.
 
@@
 #| label: fig-2
-#| fig-cap: "Scatterplot depicting the **Wins** vs the **3 Points Made** per game"
+#| fig-cap: "Scatterplot depicting the **Wins** vs the **3 Points Made** per game by all teams across the seasons 2014-15 to 2022-23"
 #| warning: false
@@
 #| label: tbl-3
-#| tbl-cap: "Table showing the win stats and 3 point stats grouped by the Team"
+#| tbl-cap: "Summary of Team statistics across seasons 2014-15 through 2022-23"
 #| warning: false
@@
 #| label: fig-3
-#| fig-cap: "Graph showing the average 3 point shots made per game across all the seasons"
+#| fig-cap: "Visualization of Average 3-Pointers Made per Game Across NBA Seasons"
 #| warning: false
@@
 #| label: fig-1
-#| fig-cap: "Graph depicting the **Average team wins** and **predicted team wins** vs **Average three point shots made** per game"
+#| fig-cap: "Relationship Between **Average Team Wins**, **Predicted Wins**, and **Average Three-Point Shots Made per Game**"
 #| warning: false
@@
 
-Although, this paper discusses the upside of the three point line, I feel it is important to also include it's drawbacks.
+Although this paper discusses the upside of the three point line, I feel it is important to also include its drawbacks.
 An interesting example of this in the modern NBA is the 2017-18 Houston Rockets.

### 2024-04-19 paper/paper.qmd
Update paper.qmd

@@
 This study adds to the ongoing debate in the NBA about strategic decision-making and player development by explaining the significance of three-point shooting in deciding game outcomes (@cite-debate).
-Essentially, the estmand of the paper is to see if an increase in a team's three point shooting ability contributes to an increase in the team's success, measured by wins.
+Essentially, the estimand of this paper is analysing if an increase in a team’s three point shooting ability contributes to an increase in the team’s success, measured by wins.
 
@@
 
-## Variables of interest
+## Variables of Interest
 
@@
 It indicates the variability in team success (measured by the number of wins) that is not explained by the number of 3-point shots made.
-The **Mean_PPD (40.3)**represents the average posterior predictive distribution of the outcome variable (number of wins).
+The **Mean_PPD (40.3)** represents the average posterior predictive distribution of the outcome variable (number of wins).
 It provides an estimate of the average number of wins predicted by the model.
@@
 This team was a huge success in the regular season and they shot more 3 pointers than any other team in league history.
-What he failed to take into account was, that basketball isn't math.
+What he failed to take into account was that basketball isn't math.
 While the Rockets were the most successful regular season team with 65 wins (@cite-Rockets2), they matched up against the Golden State Warriors in the western conference finals of the playoffs.
